---
title: "Postionen Parteien"
output:
  html_document:
    code_folding: hide
    fig_width: 6
    toc: yes
  pdf_document:
    toc: yes
---
<style type="text/css"> <!-- .table { width: auto } ---> </style>


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(knitr)
```

# GLES Wahlstudie

German Longitudinal Election Study (GLES) 2017 -- [gles.eu](http://gles.eu)

```{r, message=FALSE, warning=FALSE}
gles_raw <- haven::read_dta("daten/gles-2017/ZA6801_v2-0-0.dta", encoding = "latin1")
```

## Q-32: Links-Rechts-Selbsteinstufung

_Fragetext:_
Und wie ist das mit Ihnen selbst? Wo würden Sie sich auf der Skala von 1 bis 11 einordnen?


```{r, message=FALSE, warning=FALSE, fig.height=2.5, fig.width=3.5}
lr_waehler <- 
  gles_raw %>% 
  mutate(
    variable = "waehler_selbst",
    partei = as_factor(q19ba) %>% as.character(), # zweitstimme
    links_rechts = ifelse(q32 < 0, NA, q32)
    ) %>% 
  filter(str_detect(partei, ("^[A-Z]"))) %>% 
  select(variable, partei, links_rechts)

ggplot(lr_waehler, aes(x = links_rechts)) +
  geom_density(adjust=5)
```

<p style="page-break-before: always">

## Q-31: Links-Rechts-Einstufung Parteien

_Fragetext:_
In der Politik reden die Leute häufig von "links" und "rechts". Wenn Sie diese Skala von 1 bis 11
benutzen, wo würden Sie die folgenden Parteien einordnen, wenn 1 "links" und 11 "rechts" ist? 

Items: (A) CDU; (B) CSU; (C) SPD; (D) DIE LINKE; (E) GRÜNE; (F) FDP; (G) AfD

```{r, message=FALSE, warning=FALSE}
q31_partei <-
  tibble(
    q31 = letters[1:7],
    partei = c("CDU/CSU", "CDU/CSU", "SPD", "DIE LINKE", "GRUENE", "FDP", "AfD")
  )

lr_q31 <- 
  gles_raw %>% 
  mutate(zweitstimme = as_factor(q19ba) %>% as.character()) %>% 
  select(zweitstimme, starts_with("q31")) %>% 
  gather(q31, links_rechts, -zweitstimme)

lr_partei <- 
  lr_q31 %>%  
  mutate(
    variable = "befragte_partei",
    q31 = str_replace(q31, "q31", "")
    ) %>%
  left_join(q31_partei) %>% 
  filter(links_rechts >= 0) %>% 
  select(-q31)

lr_wa_pa <- 
  lr_partei %>% 
  filter(zweitstimme == partei) %>% 
  mutate(variable = "waehler_partei")
```

#### Positionen Parteien

```{r, message=FALSE, warning=FALSE}
partei_factor <- c("DIE LINKE", "GRUENE", "SPD", "FDP", "CDU/CSU", "AfD")
lr <- 
  bind_rows(lr_waehler, lr_partei, lr_wa_pa) %>% 
  mutate(partei = fct_relevel(partei, partei_factor))

ggplot(lr, aes(x = links_rechts, colour = variable)) +
  geom_density(adjust=5) +
  facet_wrap(~partei)
```

<p style="page-break-before: always">

# CHES expert survey

https://www.chesdata.eu/1999-2014-chapel-hill-expert-survey-ches-trend-file

## Two-dimensional space


```{r}
ches_raw <- read_dta("daten/ches-2014/1999-2014_CHES_dataset_means-3.dta")

ches_deu <-
  ches_raw %>%
  zap_labels() %>% 
  filter(country == 3, vote > 4) %>%
  mutate(
    year = as.factor(year),
    party = case_when(
      party %in% c("LINKE", "Linkspartei/PDS", "PDS") ~ "Linke",
      party == "Grunen" ~ "Grüne",
      TRUE ~ party
      )
    )

ggplot(ches_deu, aes(x=lrecon, y=galtan, colour = party, shape = year)) + geom_point()

```

__LRECON__ -- position of the party in YEAR in terms of its ideological __stance on economic issues__.
Parties can be classified in terms of their stance on economic issues. Parties on the economic left
want government to play an active role in the economy. Parties on the economic right emphasize a
reduced economic role for government: privatization, lower taxes, less regulation, less government
spending, and a leaner welfare state.

__GALTAN__ -- position of the party in YEAR in terms of of their views on __democratic freedoms
and rights__. “Libertarian” or “postmaterialist” parties favor expanded personal freedoms, for example,
access to abortion, active euthanasia, same-sex marriage, or greater democratic participation.
“Traditional” or “authoritarian” parties often reject these ideas; they value order, tradition, and
stability,

<p style="page-break-before: always">

## Most important issue in 2014

Most Important Issue -- Only asked in 2014

Entries for the next three questions are a summary of the expert responses to the Most Important
Issue question. Each expert ranked one issue as the most important, one issue as the second most
important, and one issue as the third most important issue. In this dataset, we aggregate these
responses using a simple ordinal voting technique. For each party, an issue received 10 points if it
is ranked as the #1 issue by an expert, 5 points if it is ranked #2 by an expert, and 1 point if it is
ranked #3 by an expert. After adding together the issue scores for all the experts for each individual
party, we ranked each issue by the total number of points, yielding the MIP_ONE, MIP_TWO,
and MIP_THREE variables. As the table shows, some parties have ties for the most important
issue questions, which reflects the experts’ uncertainty or disagreement on particular parties.11

 * MIP_ONE = most important issue for the party over the course of YEAR.
 * MIP_TWO = second most important issue for the party over the course of YEAR.
 * MIP_THREE = third most important issue for the party over the course of YEAR.

```{r, message=FALSE, warning=FALSE}
mip_issues <- read_csv("daten/ches-2014/most-important-issue.csv")
mip <-
  ches_deu %>% 
  select(party, year, starts_with("mip_")) %>% 
  gather(key, id, -party, -year) %>% 
  filter(! is.na(id)) %>% 
  left_join(mip_issues) %>% 
  arrange(party, key) %>% 
  select(party, key, issue)

mip %>%
  spread(key, issue) %>%
  select(party, mip_one, mip_two, mip_three) %>% 
  kable()
```

<p style="page-break-before: always">

# Manifesto content analysis

https://manifesto-project.wzb.eu/

## rile -- right-left position of party

Right-left position of party as given in Michael Laver/Ian Budge (eds.):
Party Policy and Government Coalitions, Houndmills, Basingstoke, Hampshire: The MacMillan Press 1992:

```
rile = 
  (per104 + per201 + per203 + per305 + per401 + per402 + per407 + 
   per414 + per505 + per601 + per603 + per605 + per606)
  - (per103 + per105 + per106 + per107 + per403 + per404 + per406 + 
     per412 + per413 + per504 + per506 + per701 + per202)
```

```{r, message=FALSE, warning=FALSE}
manifesto_raw <- read_csv("daten/manifesto-2017//MPDataset_MPDS2017b.csv")

ma_deu <- 
  manifesto_raw %>% 
  filter(countryname == "Germany", pervote > 4.5, date > 196000) %>% 
  mutate(
    year = date %/% 100,
    rile = round(rile, 1),
    Partei = case_when(
      party %in% c(41111, 41113) ~ "B90/Grüne",
      party %in% c(41221, 41222, 41223) ~ "Linke/PDS",
      TRUE ~ partyabbrev
    )
  )

ggplot(ma_deu, aes(x = year, y = rile, colour = Partei, linetype = Partei)) + geom_line()
```

<p style="page-break-before: always">

## Germany 2017 manifestos

```{r, message=FALSE, warning=FALSE}
ma_var <- read_csv("daten/manifesto-2017/variables.csv")

path <- "daten/manifesto-2017/Germany-2017/"
ma_files <- list.files(path, pattern = "csv$")
ma_2017_raw <- map_df(ma_files, ~ read_csv(paste0(path, .x)) %>% mutate(file = .x))

ma_2017 <- 
  ma_2017_raw %>% 
  mutate(party = str_extract(file, "\\d+") %>% as.integer()) %>% 
  left_join(ma_deu %>% distinct(party, partyabbrev)) %>% 
  filter(! cmp_code %in% c("H", "0")) %>% 
  select(partyabbrev, cmp_code, content) %>% 
  group_by(partyabbrev)
```

#### Most frequent categories

```{r, message=FALSE, warning=FALSE}
cmp_code_info <- 
  ma_var %>%
  select(cmp_code = variable, label) %>% 
  mutate(cmp_code = str_remove(cmp_code, "^per"), 
         cmp_code = str_replace_all(cmp_code, "_", "."))

ma_2017 %>%
  count(cmp_code) %>% 
  arrange(-n) %>% 
  left_join(cmp_code_info) %>% 
  slice(1:3) %>% 
  kable
```

#### Examples classified quasi-sentences

```{r, message=FALSE, warning=FALSE}
ma_2017 %>% 
  left_join(cmp_code_info) %>% 
  slice(1:4) %>% 
  kable
```

