---
title: "Manifestos · Moral Values"
author: "Holger Doering --- doering@uni-bremen.de"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
  html_notebook:
    code_folding: hide
---
<style type="text/css"> <!-- .table { width: auto } ---> </style>

```{r options, include=FALSE}
knitr::opts_chunk$set(
  # results="hide",
  message=FALSE,
  warning=FALSE,
  package.startup.message = FALSE
  )

options(
  readr.num_columns = 0,
  knitr.kable.NA = "",
  width = 100,
  tidyverse.quiet = TRUE
)
```

<!-- <p style="page-break-before: always"> -->


#### Manifesto Project Dataset (version 2018b)

* Volkens, Andrea / Krause, Werner / Lehmann, Pola / Matthieß, Theres / Merz, Nicolas / Regel, Sven / Weßels, Bernhard (2018): The Manifesto Data Collection. Manifesto Project (MRG/CMP/MARPOR). Version 2018b. Berlin: Wissenschaftszentrum Berlin für Sozialforschung (WZB). https://doi.org/10.25522/manifesto.mpds.2018b
* [manifesto-project.wzb.eu](https://manifesto-project.wzb.eu/datasets)


__per603__ -- Traditional Morality: Positive

Favourable mentions of traditional and/or religious moral values. May include:

* Prohibition, censorship and suppression of immorality and unseemly behaviour;
* Maintenance and stability of the traditional family as a value;
* Support for the role of religious institutions in state and society.

__per604__ -- Traditional Morality: Negative

Opposition to traditional and/or religious moral values. May include:

* Support for divorce, abortion etc.;
* General support for modern family composition;
* Calls for the separation of church and state

```{r corepacks}
library(tidyverse)
library(glue)
```

#### Europe

```{r}
mp_raw <- read_csv("MPDataset_MPDS2018b.csv")
```


```{r}
country_select <- c("Denmark", "Germany", "Netherlands", "United Kingdom")

mp <- mp_raw %>% 
  mutate(decade = round(date / 1000) * 10) %>% 
  filter(countryname %in% country_select)
```


```{r, eval=FALSE}
pl_dt <- 
  mp %>% 
  mutate(morality = per603 + per604) %>% 
  group_by(countryname, decade) %>% 
  summarise(morality = mean(morality))

ggplot(pl_dt, aes(x = decade, y = morality)) +
  geom_line() +
  facet_wrap(~ countryname)
```

```{r}
pl_dt <- 
  mp %>% 
  group_by(countryname, decade) %>% 
  summarise_at(c("per603", "per604"), mean) %>% 
  gather(variable, share, -countryname, -decade)

ggplot(pl_dt, aes(x = decade, y = share, colour = variable)) +
  geom_line() +
  facet_wrap(~ countryname)
```

#### Germany

```{r}
party_select <- c("CDU/CSU", "SPD", "FDP", "Greens")

pl_dt <- 
  mp_raw %>% 
  mutate(partyabbrev = ifelse(party %in% c(41111, 41112, 41113),
                              "Greens",
                              partyabbrev)) %>% 
  filter(
    countryname == "Germany",
    partyabbrev %in% party_select
    ) %>% 
  mutate(
    year = floor(date / 100),
    partyabbrev = fct_relevel(partyabbrev, party_select)
    ) %>% 
  select(year, party=partyabbrev, per603, per604) %>% 
  gather(variable, share, -year, -party)

ggplot(pl_dt, aes(x = year, y = share, colour = variable)) +
  geom_line() +
  facet_wrap(~ party)
```


#### Corpus examples

```{r}
mp_partyname <- 
  mp_raw %>% 
  select(party, partyname) %>% 
  distinct(party, .keep_all = TRUE)

cp_all <- 
  read_csv("mp-corpus/mp-morality.csv") %>% 
  left_join(mp_partyname)
```


```{r, eval=FALSE}
# examples for exercise campaign posters
cp_all %>% 
  filter(country %in% c("Germany", "United Kingdom")) %>% 
  group_by(country, code) %>% 
  sample_n(5) %>% 
  arrange(year) %>% 
  mutate(example = glue("{statement} --- {year}, {partyabbrev}, {code}")) %>% 
  pull(example)
```


```{r}
cp_all %>% 
  group_by(country, partyname) %>% 
  summarise(n_entries = n()) %>% 
  DT::datatable()

cp_all %>% DT::datatable()
```

